\batcave: Calling somatic mutations with a tumor- and site-specific prior

### 2019-12-09 NAR_DRAFT.tex
fix watsoncrick

\subsection{Estimation of the mutation profile}
Many aspects of genomic architecture can affect the somatic mutation rate at multiple scales \cite{Buisson2019}.
Here we focus on a small-scale feature, the trinucleotide context, which is known to strongly affect the prior probability of single-nucleotide mutation \cite{Nik-Zainal2012a,Alexandrov2015,Lee-Six2018}.
The trinucleotide context of a genomic site consists of the identity of the reference base and the 3' and 5' flanking bases.
\bkmcomment{There are six classes of base substitution -- C$>$A, C$>$G, C$>$T, T$>$A, T$>$C, T$>$G (all substitutions are referred to by the pyrimidine of the mutated Watson–Crick base pair).
This results in a total of 32 ($4\times 2 \times 4$) reference contexts and 3 alternate bases per reference base.}
%Folding the central base to the pyrimidines, there are two possible bases at the focal site, and there are four possible bases 3' and 5' of the focal site, yielding $2 \cdot 4 \cdot 4$ possible tri-nucleotide contexts $C$.
%At the focal site, a mutation $m$ can be to any of three alternate alleles.
Indexing by the $c=\{1 \dots 32\}$ contexts and by the $m = \{1 \dots 3\}$ alternate bases, we have 96 possible substitution types $S_{m,c}$.
Eq.~\ref{eqn:full_prior} is then
\begin{equation}
  \label{eqn:detailed_prior}
  \mathrm{p}(\mathrm{S}_{m,c}) = \mathrm{p}(m \mid C = c) \mathrm{p}(C = c \mid M)\frac{\mathrm{p}(M)}{\mathrm{p}(C = c)}.
\end{equation}
The first two terms on the right-hand side can be estimated from the observed mutation profile (Fig.~\ref{NAR-realsigfig}).
%Here $\mathrm{p}(\mathrm{S}_{m,c})$ represents probability that a substitution of type $\mathrm{S}_{m,c}$ will occur.

>>>>

\subsection{Estimation of the mutation profile}
Many aspects of genomic architecture can affect the somatic mutation rate at multiple scales \cite{Buisson2019}.
Here we focus on a small-scale feature, the trinucleotide context, which is known to strongly affect the prior probability of single-nucleotide mutation \cite{Nik-Zainal2012a,Alexandrov2015,Lee-Six2018}.
The trinucleotide context of a genomic site consists of the identity of the reference base and the 3' and 5' flanking bases.
\bkmcomment{There are six classes of base substitution -- C$>$A, C$>$G, C$>$T, T$>$A, T$>$C, T$>$G (all substitutions are referred to by the pyrimidine of the mutated Watson - Crick base pair).
This results in a total of 32 ($4\times 2 \times 4$) reference contexts and 3 alternate bases per reference base.}
%Folding the central base to the pyrimidines, there are two possible bases at the focal site, and there are four possible bases 3' and 5' of the focal site, yielding $2 \cdot 4 \cdot 4$ possible tri-nucleotide contexts $C$.
%At the focal site, a mutation $m$ can be to any of three alternate alleles.
Indexing by the $c=\{1 \dots 32\}$ contexts and by the $m = \{1 \dots 3\}$ alternate bases, we have 96 possible substitution types $S_{m,c}$.
Eq.~\ref{eqn:full_prior} is then
\begin{equation}
  \label{eqn:detailed_prior}
  \mathrm{p}(\mathrm{S}_{m,c}) = \mathrm{p}(m \mid C = c) \mathrm{p}(C = c \mid M)\frac{\mathrm{p}(M)}{\mathrm{p}(C = c)}.
\end{equation}
The first two terms on the right-hand side can be estimated from the observed mutation profile (Fig.~\ref{NAR-realsigfig}).
%Here $\mathrm{p}(\mathrm{S}_{m,c})$ represents probability that a substitution of type $\mathrm{S}_{m,c}$ will occur.
